---
title: "NFL prediction"
author: "Keith E Giles"
date: "1/23/2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Prediction of straight-up winners of NFL games. 
The goal of this project is to predict the winner of NFL games.  
The dataset was downloaded from a publicly available dataset from kaggle ("https://www.kaggle.com/zynicide/nfl-football-player-stats"). The data was made available in $\textbf{JSON}$ format, and I converted it into a dataframe upon being imported into $\textbf{R}$.  

```{r import and convert raw data, eval = F}
library(jsonlite)
NFLgamesJSON <- fromJSON("/data/kegiles/ToyData/games_1512362753.8735218.json")
NFLgamesJSON <- lapply(NFLgamesJSON, function(x) {
x[sapply(x, is.null)] <- NA
unlist(x)})
tmp <- do.call("rbind", NFLgamesJSON)
tmp <- t(tmp)
tmpDF <- data.frame(tmp)
AllGames <- tmpDF
head(AllGames)
dim(AllGames)
```
This dataset has stats from over 1.02m players and 46 features, which is a mix of player stats and data about the game.  The next goal is convert this player-specific data into Team-specific data in two steps:

1.    Create a **TeamGameID** to group each teams stats for a given game.
2.    Create a **UniqueGameID** for each game(there should be only 2 per game, one per team)
3.    Decide on relevant stats
4.    Isolate the Home teams Stats
5.    Isolate the Away team stats
6.    Join these two stats using **UniqueGameID** as the unique identifier
7.    Calculate the stat differential for each game
8.    Sort by team and date, calculate rolling mean for each team.  
9.    Drop **"game_location==A"**, which is redundant with the teams swapped and differentials inverted  
10.   Use this as a training set  to: 
         a.  predict score
         b.  predict outcome

```{r feature engineering and grouping, eval= F}
#load the dplyr package to perform grouping operations
library(plyr)
library(dplyr)
AllGames[1:46] <- lapply(AllGames[1:46], function(x) as.character(x))
#create unique ID by merging together game-related features
#This unique ID has to be the same for the redundant entries of each game, with the home team first, then the date, then the away team.  
AllGames$TeamID <- paste0(AllGames$team,AllGames$date,AllGames$opponent)
AllGames[12:46] <- lapply(AllGames[12:46], function(x) as.numeric(as.character(x)))
AllGames[1:5,1:5]
colnames(AllGames)
#The columns in the AllGames dataframe
dim(AllGames)
#The dimensions of the AllGames dataframe:
```

We now have created a new dataframe **GameSummary** that has grouped all the stats from each player from each game  from each team. In order to determine the Net Stats or "stat differential", we need to join each teams stats with their opponents stats.  This requires generating a **GameUniqueID**. You'll know this was done proprely is there are two instances of each **GameUniqueID**.  
```{r create unique game ID, eval = F}
GameSummary <- group_by(AllGames[c(7,12:47)], TeamID, game_location) %>% summarise_all(funs(sum))
#First we need to join a few of the columns that were dropped when we used groupBy to create the GameSummary dataframe. These can be joined using the TeamID field.
tmpDF <- AllGames[c(3,6,8,10,11,47)]
tmpGrouped <- group_by(tmpDF, TeamID) %>% summarise_all(funs(first))
GameSummary <- merge(x = GameSummary,y = tmpGrouped,by = "TeamID")
#Now we have the total team stats for each game AND both team names and the date and score. We can now move on to make a unique game ID
GameSummary$UniqueGameID <- ifelse(GameSummary$game_location == "H", paste0(GameSummary$team, GameSummary$date,GameSummary$opponent), paste0(GameSummary$opponent,GameSummary$date,GameSummary$team))
#the above code created the UniqueGameID, which puts the home team first.
#Now we can separate into Home and Away teams
HomeTeams <- subset(GameSummary, game_location == "H")
AwayTeams <- subset(GameSummary, game_location == "A")
colnames(HomeTeams)[3:37] <- paste0("Home_",colnames(HomeTeams)[3:37])
colnames(AwayTeams)[3:37] <- paste0("Away_", colnames(AwayTeams)[3:37])
#change player and opponent to home and away, remember this needs to be inverted for the away team dataset
colnames(HomeTeams) <- gsub("player_team_score","HomeScore",colnames(HomeTeams))
colnames(HomeTeams) <- gsub("opponent_score","AwayScore",colnames(HomeTeams))
colnames(AwayTeams) <- gsub("player_team_score","AwayScore",colnames(AwayTeams))
colnames(AwayTeams) <- gsub("opponent_score","HomeScore",colnames(AwayTeams))
colnames(HomeTeams) <- gsub("team","Home",colnames(HomeTeams))
colnames(HomeTeams) <- gsub("opponent","Away",colnames(HomeTeams))
colnames(AwayTeams) <- gsub("team","Away",colnames(AwayTeams))
colnames(AwayTeams) <- gsub("opponent","Home",colnames(AwayTeams))
```
We now have two dataframes that contain the home and away teams separted.  The next step is to join them using the UniqueGameID as the primary key. Once they are joined, we subtract the Away stats from the Home stats to generate the first differential. 
```{r join home and away stats, eval = F}
#Trim away redundant columns from Away teams
AwayTeams <- AwayTeams[-c(1,2,38:42)]
Joined <- merge(x = HomeTeams, y = AwayTeams, by = "UniqueGameID") 
Joined <- Joined[order(Joined$UniqueGameID),]
tmpHome <- data.matrix(Joined[c(4:38)])
tmpAway <- data.matrix(Joined[c(44:78)])
DifferentialMatrix <- tmpHome - tmpAway
tmpColumnNames <- colnames(Joined)[4:38]
tmpColumnNames <- gsub("Home","Delta",tmpColumnNames)
#Join up the differential stats with the game identifiers 
tmpDiff <- data.frame(DifferentialMatrix)
colnames(tmpDiff) <- tmpColumnNames
tmpJoiner <- Joined[c(1,39:43)]
DifferentialDF <- cbind(tmpJoiner, tmpDiff)
#we sorted the dataset so we have each teams game in chronological order
DifferentialDF <- DifferentialDF[order(DifferentialDF$Home,DifferentialDF$date),]
DifferentialDF[5:6] <- lapply(DifferentialDF[5:6], function(x) as.numeric(x))
DifferentialDF$Delta_Score <- DifferentialDF$HomeScore - DifferentialDF$AwayScore
DifferentialDF <- subset(DifferentialDF, select = -c(UniqueGameID,Home,Away,HomeScore,AwayScore))
```

From the **DifferentialDF** dataset we now need to calculate the rolling average of stats. Obviously, the datat in its current form is not something we'll have at the outset of each game, so its of no use for predictive purposes.  We will however, have the data from the previous games.   So our strategy is to use the average from the last five games to predict the outcome
```{r rolling average calculation, eval = F}
#Calculate the 5 game rolling average for each team
DifferentialDF_MA5 <- DifferentialDF
library(zoo)
DifferentialDF_MA5[2:36] <- lapply(DifferentialDF[2:36], function(x) rollmean(x = x, k = 5, align = "right", fill = NA))
#add the actual score to regress against alongside the rolling average of score differential 
DifferentialDF_MA5 <- cbind(DifferentialDF[c(37)],DifferentialDF_MA5)
#write.table(DifferentialDF_MA5, "/home/kegiles/ToyData/PredictingNFLwinners/DifferentialStats.tsv",sep = "\t",row.names = F,col.names = T, quote = F)
colnames(DifferentialDF_MA5) <- c("DeltaActualScore",paste0(colnames(DifferentialDF_MA5)[2:38],"_MA5"))
colnames(DifferentialDF_MA5)[2:38]<- gsub("Delta_","",colnames(DifferentialDF_MA5)[2:38])
colnames(DifferentialDF_MA5)[1] <- "NetScore"
DifferentialDF_MA5 <- DifferentialDF_MA5[-c(2)]
colnames(DifferentialDF_MA5)
```

We have finally created a dataframe which has the rolling average in 35 statistics over the home teams previous 5 games and the corresponding net points in the game. And most importantly **These stats will be available at the beginning of each game, and can be used to actually predict game outcomes**. This wil be addressed in the next notebook.
